---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

To load the data from the .csv file (which is located in a subfolder `activity/`) into a variable `rawdata`, the following command is used:

```{r}
rawdata <- read.csv("activity/activity.csv")
```

After loading the data, it has to be prepared for the following analyses by removing the NA values from the data:

```{r}
procdata <- rawdata[!is.na(rawdata["steps"]),]
```



## What is mean total number of steps taken per day?

As a basis for the calculation of mean and median, the sum of steps for each day is calculated by aggregating values obtained on the same day:

```{r}
totalNumberStepsPerDay <- aggregate(procdata$steps, by=list(procdata$date), FUN=sum)
```

A histogram of can be plotted using he hist function:

```{r}
names(totalNumberStepsPerDay) <- c("Date", "Steps_total")
hist(totalNumberStepsPerDay$Steps_total, breaks=25, main="Histogram of total number of steps per day", xlab="Total number of steps per day")
```


The mean and median of the total number of steps are calculated:

```{r}
meanStepsIgnoredNAs <- mean(totalNumberStepsPerDay$Steps_total)
medianStepsIgnoredNAs <- median(totalNumberStepsPerDay$Steps_total)
```

And the results are as follows:

```{r}
meanStepsIgnoredNAs
medianStepsIgnoredNAs
```



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
